### 2014-10-07 paper/paper3.tex
ideas of Kai

% This is LLNCS.DEM the demonstration file of
% the LaTeX macro package from Springer-Verlag
% for Lecture Notes in Computer Science,
% version 2.4 for LaTeX2e as of 16. April 2010
%
\documentclass{llncs}

% allows for temporary adjustment of side margins
\usepackage{chngpage}

% just makes the table prettier (see \toprule, \bottomrule, etc. commands below)
\usepackage{booktabs}

\usepackage[utf8]{inputenc}

% URL handling
\usepackage{url}
\urlstyle{same}

% Todos
%\usepackage[colorinlistoftodos]{todonotes}
%\newcommand{\ke}[1]{\todo[size=\small, color=orange!40]{\textbf{Kai:} #1}}
%\newcommand{\tb}[1]{\todo[size=\small, color=green!40]{\textbf{Thomas:} #1}}

%\usepackage{makeidx}  % allows for indexgeneration

%\usepackage{amsmath}
\usepackage{amsmath, amssymb}
\usepackage{mathabx}

% monospace within text
\newcommand{\ms}[1]{\texttt{#1}}

% examples
\usepackage{fancyvrb}
\DefineVerbatimEnvironment{ex}{Verbatim}{numbers=left,numbersep=2mm,frame=single,fontsize=\scriptsize}

\usepackage{xspace}
% Einfache und doppelte Anfuehrungszeichen
\newcommand{\qs}{``} 
\newcommand{\qe}{''\xspace} 
\newcommand{\sqs}{`} 
\newcommand{\sqe}{'\xspace}

% checkmark
\usepackage{tikz}
\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}

% Xs
\usepackage{pifont}

% Tabellenabstände kleiner
\setlength{\intextsep}{10pt} % Vertical space above & below [h] floats
\setlength{\textfloatsep}{10pt} % Vertical space below (above) [t] ([b]) floats
% \setlength{\abovecaptionskip}{0pt}
% \setlength{\belowcaptionskip}{0pt}

\usepackage{tabularx}
\newcommand{\hr}{\hline\noalign{\smallskip}} % für die horizontalen linien in tabellen

% Todos
\usepackage[colorinlistoftodos]{todonotes}
\newcommand{\ke}[1]{\todo[size=\small, color=orange!40]{\textbf{Kai:} #1}}
\newcommand{\tb}[1]{\todo[size=\small, color=green!40]{\textbf{Thomas:} #1}}

\setcounter{secnumdepth}{5}

\begin{document}

%
%
\title{XXXXX}
%
\titlerunning{XXXXX}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{XXXXX\inst{1} \and XXXXX\inst{2}}
%
\authorrunning{XXXXX} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\institute{XXXXX\\
\email{XXXXX},\\ 
\and
XXXXX \\
\email{XXXXX} 
}

\maketitle              % typeset the title of the contribution

\begin{abstract}

\keywords{..}
\end{abstract}
%

\section{Idea}

"What I would consider the greatest success
for us would be to create something that can be mapped to any reasonable
validation method, but that makes it easy for people to define the
constraints they want to use in their application. BIBFRAME has used a
somewhat reduced DSP in this way: the AP is what drives the BIBFRAME
instance data editor. (I would love for us to have a similar editor for
the *creation* of APs.)"

That's interesting!
Kai and I had a related idea a couple of months ago ;-)

There should be 1 constraint language-independent SPIN mapping for each RDF validation constraint (requirement) 
in order to validate this constraint automatically and independently from any specific constraint language.

Then, for each constraint language (ShEx, DSP, OWL2) - we want to use in order to express our constraints, 
we need for each constraint a mapping to the language-independent constraint.
This way, we do not need a SPIN mapping for each constraint language in order to implement the validation of the language-specific constraint.
Now, we have immediately an automatic validation of these constraints.
This way, constraints expressed with different languages can be validated in the same manner without differences.

When we use a UI to define our constraints in a user-friendly way, these constraints can be either mapped to a constraint expressed in a specific constraint language 
and then to SPIN or directly to SPIN in the background.    
So, the user does not have to know how to express constraints using a constraint language such as OWL 2, DSP, or ShEx.

And if there is a new constraint not covered by any constraint language you can just define the SPIN mapping for this constraint.

\section{XX}

ontology describing constraint languages
interoperability between constraint languages
express constraint in CL A and translate constraint in CL B
provide SPARQL validation (SPIN mapping) for each general C

\section{constraint language ontology}

\section{examples}

\subsection{min cardinality}

OWL 2:
\begin{ex}
foaf:Person
    a owl:Restriction ;
		owl:minQualifiedCardinality "2"^^xsd:nonNegativeInteger ;
		owl:onProperty :hasName ;
		owl:onClass xsd:string .
\end{ex}

general constraint: min cardinality
constraint (OWL2): data min cardinality

translation to SPARQL:
\begin{ex}

\end{ex}

\subsection{object property range}

DSCL constraint (OWL 2):
\begin{ex}
:hasDog rdfs:range :Dog . 
\end{ex}

DSCL constraint (DSP):
\begin{ex}
:hasDogRange
        a dsp:DescriptionTemplate ; 
        dsp:resourceClass owl:Thing ; 
        dsp:statementTemplate [
            a dsp:NonLiteralStatementTemplate ;
            dsp:property :hasDog ; 
            dsp:nonLiteralConstraint [ 
                a dsp:NonLiteralConstraint ;
                dsp:valueClass :Dog ] ] .
\end{ex}

general constraint:
\begin{ex}
clo:objectPropertRange
    a clo:GeneralConstraint ;
		clo:requirement clo:R-28-OBJECT-PROPERTY-RANGE ;
		clo:DSCLConstraints ( clo:objectPropertRange_OWL2 clo:objectPropertRange_DSP ) ;
		clo:objectProperty :hasDog ;
		clo:range ( :Dog ) ;
\end{ex}

valid data:
\begin{ex}
:Peter 
    :hasDog :Brian .
:Brian 
    a :Dog ;     
    a owl:Thing . 
\end{ex}

invalid data:
\begin{ex}
:Peter 
    :hasDog :Brian .
:Brian 
#   a :Dog ; # commented --> constraint violation
    a owl:Thing . 
\end{ex}

general constraint: object property range
constraint (OWL2): object property range

general C is translated to SPARQL for automatic validation

for each DSCL constraint you have to define the binding to the general constraint.

binding (DSCL constraint (OWL2) - general constraint):
\begin{ex}
# OWL 2 constraint
# -----
?OPE rdfs:range ?CE .

clo:objectProperty => ?OPE
clo:range => ?CE 
\end{ex}

binding (DSCL constraint (OWL2) - general constraint):
\begin{ex}
# DSP constraint
# -----
:hasDogRange
        a dsp:DescriptionTemplate ; 
        dsp:resourceClass owl:Thing ; 
        dsp:statementTemplate [
            a dsp:NonLiteralStatementTemplate ;
            dsp:property :hasDog ; 
            dsp:nonLiteralConstraint [ 
                a dsp:NonLiteralConstraint ;
                dsp:valueClass :Dog ] ] .

# SPIN
# -----
?descriptionTemplate a dsp:DescriptionTemplate .
?descriptionTemplate dsp:resourceClass ?resourceClass .
?descriptionTemplate dsp:statementTemplate ?statementTemplate .
?statementTemplate dsp:property ?property .
?statementTemplate dsp:nonLiteralConstraint ?nonLiteralConstraint .
?nonLiteralConstraint dsp:valueClass ?valueClass .				
FILTER EXISTS { ?nonLiteralConstraint dsp:valueClass ?valueClass . }
FILTER NOT EXISTS { ?object a ?valueClass . ?nonLiteralConstraint dsp:valueClass ?valueClass . }

# binding (DSCL constraint (OWL2) - general constraint)
# -----
clo:objectProperty => ?OPE
clo:range => ?CE 
\end{ex}

binding using SPIN

SPARQL (validation of general constraint):
\begin{ex}
?x ?clo:objectProperty ?this .
?this rdf:type owl:Thing . 
FILTER NOT EXISTS { ?this a ?clo:range } .
\end{ex}

SPARQL (validation of DSCL constraint (OWL2)):
\begin{ex}
# constraint
# -----
?OPE rdfs:range ?CE .

# data 
# ----- 
?x ?OPE ?this .
?this rdf:type owl:Thing . 
FILTER NOT EXISTS { ?this rdf:type ?CE } .
\end{ex}

SPARQL (validation of DSCL constraint (DSP)):
\begin{ex}
# data 
# -----
?this a ?resourceClass .
?this ?property ?object .
FILTER ( ?property != rdf:type ) .

# constraint
# -----
?descriptionTemplate a dsp:DescriptionTemplate .
?descriptionTemplate dsp:resourceClass ?resourceClass .
?descriptionTemplate dsp:statementTemplate ?statementTemplate .
?statementTemplate dsp:property ?property .
?statementTemplate dsp:nonLiteralConstraint ?nonLiteralConstraint .
?nonLiteralConstraint dsp:valueClass ?valueClass .				
FILTER EXISTS { ?nonLiteralConstraint dsp:valueClass ?valueClass . }
FILTER NOT EXISTS { ?object a ?valueClass . ?nonLiteralConstraint dsp:valueClass ?valueClass . }
\end{ex}

ontology design pattern.org
constraint grammar
constraint elements are the same
DSP uses other design pattern / 
constraint / constraint elements / constraint design patterns / constraint language
define terminology
related to requirements DB / relate requirements to constraints, constraint languages
constraint element: e.g. min cardinality
dependency between requirements / when this requirements is fulfilled then this is also fulfilled (e.g. min card and requ. car)
min car more powerful than req. propery
why ontology? automatic translation from DSP to OWL 
define general constraint language
we do not invent a new constraint languages
requirements in SPARQL umsetzten
only one validation for each requirement
constraints may be composed

\bibliography{literatur}{}
\bibliographystyle{plain}
\setcounter{tocdepth}{1}
%\listoftodos
\end{document}